### 2025-08-27 download-football-data.qmd
20250827 - references

Below, we provide examples for how to download various types of NFL data using the `nflreadr` [@R-nflreadr] and `nflfastR` [@R-nflfastR] packages.
For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
We save each data file after downloading it, so we can use the data in subsequent chapters.
If you have difficulty downloading the data files using the `nflreadr` [@R-nflreadr] or `nflfastR` [@R-nflfastR] packages, we also saved the data files so they are publicly available on the Open Science Framework: <https://osf.io/z6pg4>.

>>>>

Below, we provide examples for how to download various types of NFL data using the `nflreadr` [@R-nflreadr] and `nflfastR` [@R-nflfastR] packages that are part of the `nflverse` ecosystem of packages [@R-nflverse].
For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
We save each data file after downloading it, so we can use the data in subsequent chapters.
If you have difficulty downloading the data files using the `nflreadr` [@R-nflreadr] or `nflfastR` [@R-nflfastR] packages, we also saved the data files so they are publicly available on the Open Science Framework: <https://osf.io/z6pg4>.

### 2025-08-27 getting-started.qmd
20250827 - references

- Intro to `R`: <https://www.statmethods.net>
- Video training courses in `R` skills: <https://www.pluralsight.com/search?q=R>
- `R` for Data Science: <https://r4ds.hadley.nz>
- Browse the `Cookbook for R` to find solutions to common tasks and problems: <http://www.cookbook-r.com>
- Browse the `R Graph Gallery` to find examples of various graphs: <https://r-graph-gallery.com>
- Free `Codeacademy` course on `R`: <https://www.codecademy.com/learn/learn-r>
- Free `Coursera` courses on `R`: <https://www.coursera.org/search?query=R>
- Watch these videos from `Coursera`: <https://blog.revolutionanalytics.com/2012/12/coursera-videos.html>
- `Posit`/`Rstudio` Webinars: <https://posit.co/resources/videos/>
- UCLA Stats Website: <https://stats.idre.ucla.edu/r/>
- Introduction to `R` course on `Datacamp`: <https://www.datacamp.com/courses/free-introduction-to-r>
- Learning Statistics with `R`: <https://learningstatisticswithr.com>
- Teaching `R` in a Kinder, Gentler, More Effective Manner: <https://github.com/matloff/TidyverseSkeptic>
- Learn `R` interactively with `swirl`: <https://swirlstats.com>
- Use the `learnr` package [@R-learnr]: <https://rstudio.github.io/learnr/>
- Resources for learning `tidyverse` [@R-tidyverse; @Wickham2019_packages], which is a collection of `R` packages for data management: <https://www.tidyverse.org/learn/>
- You will sometimes find relevant articles on `R-bloggers`: <https://www.r-bloggers.com>

@@

For instance, you can use the following code to install the `tidyverse` package [@R-tidyverse]:

@@

In base `R`, you can perform piping with the ``base::`|>` `` expression.
In `tidyverse`, you can perform piping with the ``magrittr::`%>%` `` expression.

@@

In `tidyverse`, you cans select particular columns using the `dplyr::select()` function.

@@

In `tidyverse`, you cans select particular columns using the `dplyr::filter()` function.

@@

In `tidyverse`, you cans select particular rows and columns using the `dplyr::filter()` and `dplyr::select()` functions, respectively.

@@

Below, we transform the data from long form to wide form.
The long data are in "`player`-`season`-`week` form".
We widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable) using the `tidyr::pivot_wider()` function:

@@

Conversely, we can also restructure data from wide to long using the `tidyr::pivot_longer()` function.
Here are the data in long form, after they have been transformed from wide form using `tidyverse`:

>>>>

- Intro to `R`: <https://www.statmethods.net>
- Video training courses in `R` skills: <https://www.pluralsight.com/search?q=R>
- `R for Data Science` book: <https://r4ds.hadley.nz>
    - Tutorials from the book are available via the `r4ds.tutorials` package [@R-r4ds.tutorials]: <https://ppbds.github.io/r4ds.tutorials/>
- Browse the `Cookbook for R` to find solutions to common tasks and problems: <http://www.cookbook-r.com>
- Browse the `R Graph Gallery` to find examples of various graphs: <https://r-graph-gallery.com>
- Free `Codeacademy` course on `R`: <https://www.codecademy.com/learn/learn-r>
- Free `Coursera` courses on `R`: <https://www.coursera.org/search?query=R>
- Watch these videos from `Coursera`: <https://blog.revolutionanalytics.com/2012/12/coursera-videos.html>
- `Posit`/`Rstudio` Webinars: <https://posit.co/resources/videos/>
- UCLA Stats Website: <https://stats.idre.ucla.edu/r/>
- Introduction to `R` course on `Datacamp`: <https://www.datacamp.com/courses/free-introduction-to-r>
- Learning Statistics with `R`: <https://learningstatisticswithr.com>
- Teaching `R` in a Kinder, Gentler, More Effective Manner: <https://github.com/matloff/TidyverseSkeptic>
- Learn `R` interactively with `swirl`: <https://swirlstats.com>
- Use the `learnr` package [@R-learnr]: <https://rstudio.github.io/learnr/>
- Resources for learning `tidyverse` [@R-tidyverse; @Wickham2019_packages], which is a collection of `R` packages for data management: <https://www.tidyverse.org/learn/>
- You will sometimes find relevant articles on `R-bloggers`: <https://www.r-bloggers.com>

@@

For instance, you can use the following code to install the `tidyverse` package [@R-tidyverse; @Wickham2019_packages]:

@@

In base `R`, you can perform piping with the ``base::`|>` `` expression.
In the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages], you can perform piping with the ``magrittr::`%>%` `` expression.

@@

In the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages], you can select particular columns using the `dplyr::select()` function.

@@

In the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages], you can select particular columns using the `dplyr::filter()` function.

@@

In the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages], you can select particular rows and columns using the `dplyr::filter()` and `dplyr::select()` functions, respectively.

@@

Below, we transform the data from long form to wide form.
The long data are in "`player`-`season`-`week` form".
We widen the data by two variables (`season` and `week`), using the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages], so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable) using the `tidyr::pivot_wider()` function:

@@

Conversely, we can also restructure data from wide to long using the `tidyr::pivot_longer()` function.
Here are the data in long form, after they have been transformed from wide form using the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages]:

### 2025-08-27 index.qmd
20250827 - references

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "patchwork", 
  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
  "bestNormalize", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", 
  "lavaan", "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", 
  "ordinal", "MASS", "data.table", "future", "future.apply", "SimDesign", 
  "domir", "GGally","Rglpk","TTR",
  "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))
  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
)

@@

This work would not be possible without the important contributions of many people.
This book was made possible by the hard work of the package developers of the `nflverse` ecosystem of R packages, who made football data available, including Sebastian Carl, Ben Baldwin, Lee Sharpe, Tan Ho, and John Edwards.
I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.
I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, Adam Peterson, Jesse Kartes, and Chris Russo, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
In particular, Andrew Tungate and Dennis Andersen helped develop the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], which scrapes fantasy football projections from the web and calculates projections averaged across sources.
Andrew Tungate also made important additions to the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] to allow computing fantasy points given projected (or actual) player statistics, which was necessary for conducting a [bootstrap simulation](#sec-simulationProjectedPointsBootstrapping) of players' projected fantasy points.
I am fortunate to work with such a talented, passionate, and creative team.
I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.
In addition, a special thanks to those who contributed to the book via comments, suggestions, and pull requests, including George Ellis ([\@MargareeMan](https://github.com/MargareeMan)).
Finally, I thank my wife who, in addition to designing the book cover, is a constant source of love and support.

>>>>

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "patchwork", 
  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
  "bestNormalize", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", 
  "lavaan", "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", 
  "ordinal", "MASS", "data.table", "future", "future.apply", "SimDesign", 
  "domir", "GGally","Rglpk","TTR",
  "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))
  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic", "r4ds.tutorials" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
)

@@

This work would not be possible without the important contributions of many people.
This book was made possible by the hard work of the package developers of the `nflverse` ecosystem of packages [@R-nflverse], who made football data available, including Sebastian Carl, Ben Baldwin, Lee Sharpe, Tan Ho, and John Edwards.
I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.
I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, Adam Peterson, Jesse Kartes, and Chris Russo, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
In particular, Andrew Tungate and Dennis Andersen helped develop the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], which scrapes fantasy football projections from the web and calculates projections averaged across sources.
Andrew Tungate also made important additions to the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] to allow computing fantasy points given projected (or actual) player statistics, which was necessary for conducting a [bootstrap simulation](#sec-simulationProjectedPointsBootstrapping) of players' projected fantasy points.
I am fortunate to work with such a talented, passionate, and creative team.
I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.
In addition, a special thanks to those who contributed to the book via comments, suggestions, and pull requests, including George Ellis ([\@MargareeMan](https://github.com/MargareeMan)).
Finally, I thank my wife who, in addition to designing the book cover, is a constant source of love and support.
